\systemnamenospace: Head Orientation Based Selection\\in Physical Spaces

### 2014-06-27 doc/uist14/submission/rebuttal.txt

We thank the reviewers for their detailed feedback. Below, we answer the
raised questions and clarify some points of confusion.

CONTRIBUTION  LEVEL (AC, R1, R2, R3)
Reviewers find this work’s contribution limited and suggest focusing on
interacting with complex targets. In the past, we built interfaces to
control complex appliances including smart TVs; however, CHI reviewers on
a prior application-centric paper suggested focusing on fundamental
selection techniques instead, resulting in the abbreviated description of
applications in our paper. Some examples can be found in the APPLICATION
section; we are happy to reintroduce more material in our revision.

We focus on the fundamental question of target selection itself, which is
likely to be a problem in a future rich with smart devices. Existing
commercial solutions (primarily list selection) don’t scale well (we
envision a swarm of interactive devices as the contexts) and exhibit
naming problems when multiple identical objects are within the same
space. Our work contributes a sophisticated understanding of methods for
using head-orientation as an indication of user’s point of interest in
physical space.

LOW DISAMBIGUATION SUCCESS RATE (R1)
R1 comments that our technique barely outperforms chance. However, our
success rate of 55% in iteration 2 is much higher than random chance
given that there are at least 4 targets in the user’s field of view in
our setup (in this case, chance would be 25%).

COMPUTER VISION APPROACH (AC, R2, R3)
Reviewers suggest considering computer vision-based techniques with
markers in the environment (such as QR codes, or Bokode [Mohan] which
extends the working distance to a few meters for visual tags). We
considered this direction and found several drawbacks, including
computational complexity, low accuracy when targets are far, the absence
of real-time environment feedback with passive tags, and aesthetic issues
with QR codes. While we don’t claim our choice of IR is optimal, it has
important advantages -- it is low-cost, readily available and quite
suitable for area selection. We can weaken our claim and add explanations
of how disambiguation techniques might generalize to other methods of
implementing target identification.

SYSTEM DESCRIPTION (AC)
The AC commented that “the description of the system is very
lacking”. We provided brief implementation information in order to
focus on the design and evaluation within a shorter paper to match the
magnitude of contribution. This may have been counterproductive. We list
part numbers in the "Hardware" subsection but can describe how IDs are
exchanged over IR and 802.15.4 in more detail.

MISSING RELATED WORK (AC, R3)
We will add the additional references.
RFIG Lamps and photosensing wireless tags [Raskar et al] use a projector
and data is encoded in each pixel so that tags can localize themselves
for target selection. Our IR emitter, a relatively low-cost solution,
doesn't have this feature but shares the commonality of covering an area
for the ease of selection. At the end of their paper, they envision an
IR-based system to solve ambient light problems. Our work lies in that
direction and also contributes evaluations with user studies.
Progressive Refinement [Bacima] studies several progressive refinement
selection modalities, which is close to our two-stage selection; but our
contexts are head-orientation as it reflects users’ point of interest.
Tinmith-metro [Piekarski et al] tracks the users’ hand movements and
uses a HUD for augmented reality, while our work focuses on direct
selection on physical targets.

MISC (AC)
The AC asked why users couldn't swipe in a certain direction on Glass for
disambiguation. Two major factors preclude this interaction: 1) the Glass
physical form factor makes upward swipes awkward to perform; 2) the plane
of the touchpad is orthogonal to the plane of the targets, leading to
ambiguity what lateral swipes mean.

### 2014-06-27 doc/sui14/applications.tex
edits

\sean{here are some parts I can think of to add to this section, feel free to refine/elaborate/remove.}

\changes{During our study, we learned a few things. (1)Our soulution works best on simple applications such as on/off or volume up/down controls. For more complicated controls, we suggest to use a hybrid solution in which HOBS completes the target selection process and seamlessly transfer information to user's smartphone and triggers corresponding app.  (2)Our solution excels at tasks where the user needs to focus primarily on the objects and the information provided by the near-eye display is only supportive, for example, museums or automobile body shops. (3)Our soulution also works well for roles whose hands are constantly occupied from using handheld devices, e.g. chefs.}
%% I feel like we should mention a bit about the interaction complexity. But that's a bit out of the scope of this paper.
%% Participants rated the ease of control of particular appliances differently. Ease of use ratings were higher for the lamp and fan which had simple, discrete on/off actions, and lower for the more complex movie player (see Figure~\ref{fig:smarthome-likert}). Multiple participants remarked that the difficulty was based on the affordances of Glass: \studyquote{Most of the difficulty I had with Glass came from having to navigate the interface on the tiny screen with the touch pad}. The screen size and (largely) 1D input put a limit on the complexity of interfaces that can be presented. As one participant remarked: \studyquote{[The media player] does not seem to be more efficient than a tablet device.} The difficulty can also partly be ascribed to our interaction design, which required one finger swipes to switch between parameters and two finger gestures for adjusting parameters --- it was hard for users to exert fine control over two-finger swipes. In addition, users did not always remember these mappings as they are not yet part of a standard gesture vocabulary.

>>>>

\sean{here are some parts I can think of to add to this section, feel free to refine/elaborate/remove.} 
\bjoern{1 is problematic - because it sounds like that the head worn device is in the end not worth it and you should just use your phone, even if it takes longer. I'm not sure we can confidently say something about museums or body shops - this sounds speculative. Needs section needs to be rephrased., e.g. as in: Our solution succeeds at X because of Y. This suggests that future applications in domains like Z could be especially promising.}
\changes{During our study, we learned a few things. (1)Our soulution works best on simple applications such as on/off or volume up/down controls. For more complicated controls, we suggest to use a hybrid solution in which HOBS completes the target selection process and seamlessly transfer information to user's smartphone and triggers corresponding app.  (2)Our solution excels at tasks where the user needs to focus primarily on the objects and the information provided by the near-eye display is only supportive, for example, museums or automobile body shops. (3)Our soulution also works well for roles whose hands are constantly occupied from using handheld devices, e.g. chefs.}
%% I feel like we should mention a bit about the interaction complexity. But that's a bit out of the scope of this paper.
%% Participants rated the ease of control of particular appliances differently. Ease of use ratings were higher for the lamp and fan which had simple, discrete on/off actions, and lower for the more complex movie player (see Figure~\ref{fig:smarthome-likert}). Multiple participants remarked that the difficulty was based on the affordances of Glass: \studyquote{Most of the difficulty I had with Glass came from having to navigate the interface on the tiny screen with the touch pad}. The screen size and (largely) 1D input put a limit on the complexity of interfaces that can be presented. As one participant remarked: \studyquote{[The media player] does not seem to be more efficient than a tablet device.} The difficulty can also partly be ascribed to our interaction design, which required one finger swipes to switch between parameters and two finger gestures for adjusting parameters --- it was hard for users to exert fine control over two-finger swipes. In addition, users did not always remember these mappings as they are not yet part of a standard gesture vocabulary.

### 2014-06-27 doc/sui14/interaction.tex
edits

\subsection{Hardware}
A central hypothesis of this paper is that the area cursor paradigm~\cite{kabbash1995prince} is well matched to head orientation selection. Head orientation input is imprecise because it does not capture eye movement and has to rely on a low-bandwidth muscle group. Therefore, point selection techniques like laser pointers are not appropriate.

Infrared (IR) emitters are a good technology match since they emit light within a given angle, resulting in a cone in front of the emitter where the light is visible. IR LEDs with many different fields of view are commercially available. \changes{We augment Google Glass with a 940nm 5mm IR emitter with $10^\circ$ beam angle (OSRAM SFH 4545). The emitter is controlled by an additional microcontroller which communicates with Google Glass through Bluetooth radio, since Glass does not directly support hardware modification (see Figure~\ref{fig:glass}). Target devices use Vishay TSOP38238 IR receivers. Data is encoded using standard IR remote protocols at 38.0kHz.}

\changes{IR signals are used for intial line-of-sight targeting; subsequently, we use bi-directions wireless communication to send IR signal status from targets to back to Glass and for device control itself. We have chosen a commercial off-the-shelf ZigBee implementation (XBee based on 802.15.4 radio) for this purpose (see Figure~\ref{fig:architecture}). This architecture was mostly chosen for reasons of expediency and we do not claim optimality for our design decisions. Future head-mounted devices could clearly integrate IR emitters; the choice of wireless technology could also change. In particular, one could substitute WiFi modules or design an all-Bluetooth network and forego intermediate layers in our design.}

@@

{\bf Scan:} The user first scans the environment to locate the position of the target. During this stage, Glass constantly sends out IR signals, and  targets offer immediate feedback when they receive a signal. The user confirms his desire to connect to a target by tapping on the Glass touchpad. Glass collects the responses from targets that have received IR reception through the backchannel of XBee. If there is only one single target in IR range, it is automatically selected. However, in a dense environment where multiple targets are within range, the user needs refine his selection.

>>>>

\subsection{Hardware}
%A central hypothesis of this paper is that the area cursor paradigm~\cite{kabbash1995prince} is well matched to head orientation selection. Head orientation input is imprecise because it does not capture eye movement and has to rely on a low-bandwidth muscle group. Therefore, point selection techniques like laser pointers are not appropriate.

We hypothesize that infrared (IR) emitters are a good technology match for head orientation selection, since they emit light within a given angle, resulting in a cone in front of the emitter where the light is visible. IR LEDs with many different beam angles are commercially available. \changes{We augment Google Glass with a 940nm 5mm IR emitter with $10^\circ$ beam angle (OSRAM SFH 4545). The emitter is controlled by an additional microcontroller which communicates with Google Glass through Bluetooth radio, since Glass does not directly support hardware modification (see Figure~\ref{fig:glass}). Target devices use Vishay TSOP38238 IR receivers. Data is encoded using standard IR remote protocols at 38.0kHz.}

\changes{IR signals are used for intial line-of-sight targeting; subsequently, we use bi-directions wireless communication to send information such as target IDs and signal strength from targets to back to Glass. We have chosen a commercial off-the-shelf ZigBee implementation (XBee based on 802.15.4 radio) for this purpose (see Figure~\ref{fig:architecture}). This architecture was mostly chosen for reasons of expediency and we do not claim optimality for our design decisions. Future head-mounted devices could clearly integrate IR emitters; the choice of wireless technology could also change. In particular, one could substitute WiFi modules or design an all-Bluetooth network and forego intermediate layers in our design.}

@@

{\bf Scan:} The user first scans the environment to locate the position of the target. During this stage, Glass constantly sends out IR signals, and  targets offer immediate visual feedback when they receive a signal. The user confirms his desire to connect to a target by tapping on the Glass touchpad. Glass collects the responses from targets that have received IR reception through the backchannel of XBee. If there is only one single target in IR range, it is automatically selected. However, in a dense environment where multiple targets are within range, the user needs refine his selection.

### 2014-06-27 doc/sui14/iteration1.tex
edits

Our results indicate that {\em IR selection} outperforms {\em list selection}.  The average target acquisition time for {\em IR selection} is 6.67 seconds 
while {\em list selection} took 8.86 {\em list} seconds (see Figure~\ref{fig:ir_vs_list}). A t-test shows a significant difference ($t(279)=-3.81, p=0.00017$).

@@

As for the qualitative feedback, 11 out of 14 participants preferred IR selection over list mode (two preferred list, one was undecided). While both interfaces were judged similarly on overall ease of connecting, list selection was perceived to be cumbersome. One user noted benefit of {\em IR selection} being \studyquote{more direct [than list mode]}, allowing users to focus on the targeted objects instead of the screen. \achal{This seems rather weak... Don't really need all the quotes.} One subject called it \studyquote{natural to interact with things just by looking at them}. Another mentioned that \studyquote{it's really convenient that what I'm looking at is what I'm targeting}, and many participants compared the overall time of target acquisition. One subject pointed out that \studyquote{{\em IR selection} doesn't requires you to swipe a lot of times like list mode}.

>>>>

Our results indicate that {\em IR selection} outperforms {\em list selection}.  The average target acquisition time for {\em IR selection} is 6.67 seconds 
while {\em list selection} took 8.86 seconds (see Figure~\ref{fig:ir_vs_list}). A t-test shows a significant difference ($t(279)=-3.81, p=0.00017$).

@@

As for the qualitative feedback, 11 out of 14 participants preferred IR selection over list mode (two preferred list, one was undecided). While both interfaces were judged similarly on overall ease of connecting, list selection was perceived to be cumbersome. One user noted benefit of {\em IR selection} being \studyquote{more direct [than list mode]}, allowing users to focus on the targeted objects instead of the screen. One subject called it \studyquote{natural to interact with things just by looking at them}. Another mentioned that \studyquote{it's really convenient that what I'm looking at is what I'm targeting}, and many participants compared the overall time of target acquisition. One subject pointed out that \studyquote{{\em IR selection} doesn't requires you to swipe a lot of times like list mode}.

### 2014-06-27 doc/sui14/relatedwork.tex
edits

\subsection{Head and Gaze Input}
Head movement has long been used for virtual camera control in VR applications~\cite{pausch_user_1993} and as an assistive input technology for cursor control of desktop applications~\cite{radwin1990method}. However, human neck muscles have a lower bandwidth than other muscle groups, e.g., the wrist~\cite{card_morphological_1991}.

%Card, summarizing other work, shows that neck muscles are a poor muscle group for pointing in general: neck muscles only have a bandwidth of about $4.2bits/s$ (compared to $23bits/s$ of wrist muscles used by a standard mouse~\cite{Card:1991:MAD:123078.128726}.

Prior work often focused on head orientation for controlling graphical interfaces; in contrast, we apply this modality to selection in physical spaces.

Gaze can also be used to control graphical user interfaces~\cite{kumar2007eyepoint}. While there are wearable gaze trackers~\cite{bulling2009wearable}, turning information about a concrete point in space where a user is looking into a selection requires a map of each space with known target locations. Our system works through point-to-point IR communication and does not require an {\em a priori} map. Target objects in the environment can also be equipped with individual cameras that watch the user~\cite{smith2013gaze,vertegaal2005media}. Such an approach can enable similar benefits as our approach, but is computationally more expensive than our single-pixel sensor solution and may not work at greater distances or angles, because it relies on finding the user's pupils in a camera image.

@@

\subsection{Area Cursors}
\changes{
In 2D area cursors for GUIs, the activation area of the cursor is enlarged, which facilitates acquiring smaller targets~\cite{kabbash1995prince}. We argue that head orientation pointing has analogous characteristics (limited pointing performance and accuracy). Area cursors are especially appropriate for individuals with motor control impairments or difficulties~\cite{worden1997making,findlater2010enhanced}. Similar ideas have also been extended into 3D to provide selection with progressive refinement in 3D scenes~
\cite{bacim2013design}. In all area and space cursors, the large activation area can lead to multiple targets being selected and disambiguation is needed.

We conceptualize area pointing as a two-stage process: in the {\em coarse} phase, which we call {\em scanning}, users move so the activation area intersects with the target object (and possibly other, unintended targets). In the {\em refinement} phase, they adjust so only the intended target will be selected. Many disambiguation techniques are possible for refinement -- this paper describes the trade-offs between several of them.
}

@@

Several approaches to spatial selection with handheld devices exist for controlling appliances~\cite{beigl_point_1999,patel_2-way_2003,wilson_xwand:_2003,schmidt_picontrol:_2012,kemp_point-and-click_2008} and exchanging information with smart infrastructure sensor networks ~\cite{lifton_tricorder:_2007,mittal_ubicorder:_2011,costanza_sensortune:_2010}. In some techniques, users select objects of interest with laser pointers. The laser dot provides immediate visual feedback to the user about what is being selected; however, its small target area makes it poorly matched to head orientation input.

Other approaches rely on virtual room models in which a user's location is estimated using IMU-based orientation sensing~\cite{wilson_xwand:_2003,lifton_tricorder:_2007} -- in contrast, our technique does not require a static map ahead of time.

Handheld projectors can both display a user interface in space and communicate control information optically, e.g., by encoding information temporally (using Gray codes in Picontrol~\cite{schmidt_picontrol:_2012} and RFIG Lamps~\cite{raskar_rfig_2004}) or spatially (using QR codes in the infrared spectrum in SideBySide~\cite{willis_sidebyside:_2011}). Our solution is similar in spirit but relies on simple low-cost IR emitters and detectors. \bjoern{need a stronger statement here}
%However, visible tags at appropriate sizes may be rejected by users because of their negative aesthetic effect on the space.

Other targeting systems use IR with handheld pointers~\cite{swindells_that_2002} as well as wearable devices such as rings and Bluetooth audio earpieces~\cite{merrill_augmenting_2007} to connect to smart devices. Our system tackles an unresolved issue of such IR-based approaches -- navigating an area dense with potential targets.

\subsection{Markers and Vision Methods}
\achal{Not sure where to put this, feel free to move.}
\changes{Many alternative solutions for detecting devices in contained spaces
rely on passive tags on devices, combined with computer vision methods for
detection. Unfortunately, these methods either impose significant constraints
on the camera used for detection, or require large, obtrusive tags, and
inevitably lack feedback from passive devices in the environment. \cite{Bokode},
for example, relies on visually imperceptible codes, but has significant
drawbacks: despite using a 12.1 megapixel DSLR camera, detections are no
longer robust past an angle of 20. Similarly, in our experiments, we found
that striking a balance between the size of the tag, the prices of the
device, and the computational cost of detection was unfeasible with current
cameras and algorithms.}

\ben{Reviewers suggest considering computer vision-based techniques with markers in the environment (such as QR codes, or Bokode [Mohan] which extends the working distance to a few meters for visual tags). We considered this direction and found several drawbacks, including computational complexity, low accuracy when targets are far, the absence of real-time environment feedback with passive tags, and aesthetic issues with QR codes. While we don't claim our choice of IR is optimal, it has important advantages -- it is low-cost, readily available and quite suitable for area selection. We can weaken our claim and add explanations of how disambiguation techniques might generalize to other methods of implementing target identification.}

\achal{Read and discussed the Bokode paper and CV in general. Unsure about what
to do regarding "weakening our claim."}

>>>>

\subsection{Head and Gaze Input}
\changes{Head movement has long been used for virtual camera control in VR applications~\cite{pausch_user_1993} and as an assistive input technology for cursor control of desktop applications~\cite{radwin1990method}. However, it is notable that human neck muscles have a lower bandwidth than other muscle groups, e.g., the wrist~\cite{card_morphological_1991}.}
%
%Card, summarizing other work, shows that neck muscles are a poor muscle group for pointing in general: neck muscles only have a bandwidth of about $4.2bits/s$ (compared to $23bits/s$ of wrist muscles used by a standard mouse~\cite{Card:1991:MAD:123078.128726}. 
%
Prior work often focused on head orientation for controlling graphical interfaces; in contrast, we apply this modality to selection in physical spaces.

Gaze can also be used to control graphical user interfaces~\cite{kumar2007eyepoint}. While there are wearable gaze trackers~\cite{bulling2009wearable}, turning information about a concrete point in space where a user is looking into a selection requires a map with known target locations or the use of visually obtrusive markers such as QR codes. 
Our system works through point-to-point IR communication and does not require an {\em a priori} map or markers. Target objects in the environment can also be equipped with individual cameras that watch the user~\cite{smith2013gaze,vertegaal2005media}. Such an approach can enable similar benefits as our approach, but is computationally more expensive than our single-pixel sensor solution and may not work at greater distances or angles, because it relies on finding the user's pupils in a camera image.

@@

\subsection{Area Cursors}
\changes{
In 2D area cursors for GUIs, the activation area of the cursor is enlarged, which facilitates acquiring smaller targets~\cite{kabbash1995prince}. We argue that head orientation pointing has analogous characteristics (limited pointing performance and accuracy). Area cursors are especially appropriate for individuals with motor control impairments or difficulties~\cite{worden1997making,findlater2010enhanced}. Similar ideas have also been extended into 3D to provide selection with progressive refinement in 3D scenes~
\cite{bacim2013design}. In all area and space cursors, the large activation area can lead to multiple targets being selected and disambiguation is needed. This paper describes the trade-offs between several disambiguation approaches.

%We conceptualize area pointing as a two-stage process: in the {\em coarse} phase, which we call {\em scanning}, users move so the activation area intersects with the target object (and possibly other, unintended targets). In the {\em refinement} phase, they adjust so only the intended target will be selected. Many disambiguation techniques are possible for refinement -- this paper describes the trade-offs between several of them.
}

@@

Several approaches to spatial selection with handheld devices~\cite{beigl_point_1999,patel_2-way_2003,wilson_xwand:_2003,schmidt_picontrol:_2012,kemp_point-and-click_2008} or finger-worn devices~\cite{merrill_augmenting_2007} exist. 
%and exchanging information with smart infrastructure sensor networks ~\cite{lifton_tricorder:_2007,mittal_ubicorder:_2011,costanza_sensortune:_2010}. 
In some techniques, users select objects of interest with laser pointers. The laser dot provides immediate visual feedback to the user about what is being selected; however, its small target area makes it poorly matched to head orientation input.
Other approaches rely on virtual room models in which a user's location is estimated using IMU-based orientation sensing~\cite{wilson_xwand:_2003,lifton_tricorder:_2007} -- in contrast, our technique does not require a static map ahead of time.
%Other targeting systems use IR with handheld pointers~\cite{swindells_that_2002} as well as wearable devices such as rings and Bluetooth audio earpieces~\cite{merrill_augmenting_2007} to connect to smart devices. 
Our system tackles an unresolved issue of prior approaches -- navigating an area dense with potential targets and resolving selection ambiguity.

\subsection{Vision- and Projection-Based Target Selection}
\changes{Many alternative solutions for detecting devices in contained spaces
rely on computer-vision recognition of printed tags on devices. Unfortunately, these methods either impose significant constraints
on the camera used for detection, or require large, obtrusive tags, and
inevitably lack feedback from passive markers in the environment. \bjoern{need a better reference than bokode that uses QR codes for initiating interaction.}
Handheld projectors can both display a user interface in space and communicate control information optically, e.g., by encoding information temporally (using Gray codes in Picontrol~\cite{schmidt_picontrol:_2012} and RFIG Lamps~\cite{raskar_rfig_2004}) or spatially (using QR codes in the infrared spectrum in SideBySide~\cite{willis_sidebyside:_2011}). Our solution is similar in spirit but relies on simple low-cost IR emitters and detectors. \bjoern{need a stronger statement here}}
%However, visible tags at appropriate sizes may be rejected by users because of their negative aesthetic effect on the space.

\bjoern{I tried to fold in some of the points about QR codes into the text above.}
%\subsection{Markers and Vision Methods}
%\achal{Not sure where to put this, feel free to move.}
%
%\ben{For CV, I think related work is one option to go; however, the requirement of being related work is that -- we need to reference them. There are a few examples ``require large, obtrusive tags, and inevitably lack feedback from passive devices in the environment'' which should be referenced. That's why I am thinking of putting this to discussion.}
%

%\ben{Reviewers suggest considering computer vision-based techniques with markers in the environment (such as %QR codes, or Bokode [Mohan] which extends the working distance to a few meters for visual tags). We considered this direction and found several drawbacks, including computational complexity, low accuracy when targets are far, the absence of real-time environment feedback with passive tags, and aesthetic issues with QR codes. While we don't claim our choice of IR is optimal, it has important advantages -- it is low-cost, readily available and quite suitable for area selection. We can weaken our claim and add explanations of how disambiguation techniques might generalize to other methods of implementing target identification.}

%\achal{Read and discussed the Bokode paper and CV in general. Unsure about what
%to do regarding "weakening our claim."}
